# Differentiation 

First, we take the equation as an expression

```{r}
f <- expression(x^2)
```

To calculate first derivative of $f$, we use `D()` function and `x` to specify that derivation has to be carried out with respect to 
$x$.

```{r}
f_1 <- D(f, "x")
print(f_1)
```


Sketch the graph of $f$ and $f^\prime$

```{r}
library(ggplot2)
x <- seq(-1, 1, by = 0.1)
y <- eval(f)
x <- seq(-1, 1, by = 0.1)
y1 <- eval(f_1)
data <- data.frame(x, y, y1)
head(data)

p <- ggplot(data, aes(x = x, y = y)) +
  geom_line()
print(p)
```

```{r}
q <- ggplot(data, aes(x = x, y = y1)) +
  geom_line()
print(q)
```


## Higher Derivatives

The following R command can be used to find second derivative of the above $f$.


```{r}
f_2 <- D(D(f, "x"), "x")
print(f_2)
```


```{r fig.height=3, fig.width=3}
x <- seq(-1, 1, by = 0.1)
y2 <- eval(f_2)

data <- data.frame(x, y, y1, y2)
head(data)

p <- ggplot(data, aes(x = x, y = y)) +
  geom_line()
print(p)


q <- ggplot(data, aes(x = x, y = y1)) +
  geom_line(colour = "red")
print(q)

r <- ggplot(data, aes(x = x, y = y2)) +
  geom_line(colour = "blue")
print(r)
```



## Partial Derivatives

If the expression is having more than one independent variable, we can calculate differentiation with respect to each of them.


<!-- https://www.datamentor.io/r-programming/3d-plot/-->




```{r}
f <- expression(x^2 + y^2)

x <- y <- seq(-3, 3, length = 20)
surface <- function(x, y) {
  eval(f)
}
z <- outer(x, y, surface)
persp(x, y, z)
```

Differentiate with respect to `x`

```{r}
D(f, "x")
```


Differentiate with respect to `y`

```{r}
D(f, "y")
```



