# Statistical Distributions

```{r initialstat, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.path = 'figure/',
  cache.path = 'cache/',
  fig.align = 'center',
  fig.show = 'hold',
  cache = FALSE,
  external = TRUE,
  dev = "png",
  fig.height = 5,
  fig.width = 10
)


library(tidyverse)
library(ggpubr)

```

- Density, cumulative distribution function, quantile function and random variate generation for many standard probability distributions are available in the stats package.

  - dxxx : functions for the density/mass function, 
  - pxxx : cumulative distribution function 
  - qxxx : quantile function 
  - rxxx : random variable generation.
  

## pxxx

**Cumulative distribution function (lower tail probability)**

*Example : Standard normal distribution*

- `pnorm(value-of-x-axis)` or `pnorm(quantile)`

- `pnorm(0) = 0.5` (the area under the standard normal curve to the left of zero).

```{r norm1, fig.cap="Standard normal distribution", echo=FALSE}
set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    geom_area(mapping = aes(x = ifelse(x>-50 & x< 0 , x, 0)), fill = "darkblue") +
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 0, colour ="red")+
    geom_text(aes(x=0, label="x=0", y=0.3), colour="red", angle=90, vjust = 1.2, text=element_text(size=4)) + 
   geom_segment(aes(x = 0, y = 0, xend = 0, yend = 0.2), size=1, color="green")+
  geom_segment(aes(x = 0, y = 0.2, xend = -0.7, yend = 0.2), arrow = arrow(length = unit(0.5, "cm")), size=1, color="green")

print(p)
```



- `pnorm(1.281552) = 0.9000` (the area under the standard normal curve to the left of 1.281).

```{r norm2, fig.cap="Standard normal distribution", echo=FALSE}
set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    geom_area(mapping = aes(x = ifelse(x>-50 & x< 1.281 , x, 0)), fill = "darkblue") +
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 1.281, colour ="red") +
  geom_text(aes(x=1.281, label="x=1.281", y=0.3), colour="red", angle=90, vjust = 1.2, text=element_text(size=4))+
  geom_segment(aes(x = 1.281, y = 0, xend = 1.281, yend = 0.2), size=1, color="green")+
  geom_segment(aes(x = 1.281, y = 0.2, xend = -0.2, yend = 0.2), arrow = arrow(length = unit(0.5, "cm")), size=1, color="green")

print(p)
```


- The `pnorm` function also takes the argument `lower.tail`. If `lower.tail` is set equal to `FALSE` then `pnorm` returns the upper tail probability (*the integral from $q$ to $\infty$ of the pdf*) of the normal distribution. 

- Note that 

```{r}
pnorm(1.281552)

pnorm(1.281552, lower.tail = TRUE)

pnorm(1.281552, lower.tail = FALSE) 

1-pnorm(1.281552, lower.tail = TRUE)
```

## qxxx

*Example : Standard normal distribution*

**The `qnorm` function is simply the inverse of the cdf, which you can also think of as the inverse of `pnorm`!**

- `qnorm(probability)`

- `qnorm(0.5) = 0`  (0 is the 50th percentile of the standard normal distribution)


```{r norm1b, fig.cap="Standard normal distribution", echo=FALSE}
set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    geom_area(mapping = aes(x = ifelse(x>-50 & x< 0 , x, 0)), fill = "darkblue") +
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 0, colour ="red")+
    geom_text(aes(x=0, label="x= ???", y=0.3), colour="red", angle=90, vjust = 1.2, text=element_text(size=4)) +
  geom_segment(aes(x = -0.7, y = 0.2, xend = 0, yend = 0.2), size=1, color="green")+
  geom_segment(aes(x = 0, y = 0.2, xend = 0, yend = 0), arrow = arrow(length = unit(0.5, "cm")), size=1, color="green")

print(p)
```


```{r}
qnorm(0.5)

qnorm(0.9)

qnorm(0.1, lower.tail = FALSE)

qnorm(0.9,  lower.tail = FALSE)
```


```{r norm3, fig.cap="Standard normal distribution", echo=FALSE}


cuts <- data.frame(Ref = c("x= -1.281", "x= 0", "x= 1.281"),
vals = c(-1.281, 0, 1.281),
                   stringsAsFactors = FALSE)
set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(mapping = aes(xintercept = vals, colour = Ref), data = cuts, show.legend = FALSE) +
  geom_text(mapping = aes(x = vals, y = 0.02, label = Ref, angle=90, vjust = -1.3),data = cuts)
  

print(p)
```



## dxxx


*Example : Standard normal distribution*

**The function `dnorm` returns the value of the probability density function for the  normal distribution given parameters for $x$, $\mu$, and $\sigma$.**


- `dnorm(0) == 1/sqrt(2*pi)`

If $Z\sim N(0,1)$, then

$$\phi_Z(z) = \frac{1}{\sqrt{2 \pi}}e^{-\frac{1}{2}z^2};\;\; -\infty< z< \infty$$


$$\phi_Z(0) = \frac{1}{\sqrt{2 \pi}} =  0.3989423$$


```{r}
dnorm(0)

1/sqrt(2*pi)
```


```{r norm4, fig.cap="Standard normal distribution", echo=FALSE}

set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 0, colour ="red") +
  geom_hline(yintercept = 0.3989423, colour ="red") +
  geom_text(aes(x=0, label="(0, 0.3989)", y=0.3989423), colour="blue", vjust = -1.2) +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 0.3989), arrow = arrow(length = unit(0.5, "cm")), size=1, color="green")

print(p)
```



$$\phi_Z(1) = \frac{1}{\sqrt{2 \pi}}e^{-\frac{1}{2}} =  0.3989423$$



```{r}
dnorm(1)

(1/sqrt(2*pi)) * exp(-1/2)
```



```{r norm5, fig.cap="Standard normal distribution", echo=FALSE}

set.seed(123)
data <- data.frame(x= c(rnorm(1000000,0,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    xlim(-5, 5)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 1, colour ="red") +
  geom_hline(yintercept = 0.2419707, colour ="red") +
  geom_text(aes(x=1, label="(1, 0.2419707)", y=0.2419707), colour="blue", vjust = -1.2)+
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = 0.2419707), arrow = arrow(length = unit(0.5, "cm")), size=1, color="green")

print(p)
```

## rxxx

*Example : Standard normal distribution*

`rnorm(100)` generates 100 random deviates from a standard normal distribution.

```{r}
x <- rnorm(10)
x
hist(x)
```

```{r}
y<- rnorm(1000)
hist(y)

```


## Distributions in R


Distribution | R name | Additional arguments
-------------|---------|-------------
beta         | `beta`    | `shape1, shape 2, ncp`
binomial     | `binom`   | `size, prob`
Cauchy       | `cauchy`  | `location, scale`
chi-squared  | `chisq`   | `df, ncp`
exponential  | `exp`     | `rate`
F            |  `f`      | `df1, df2, ncp`
gamma        | `gamma`   | `shape, scale`
geometric    | `geom`    | `prob`
hypergeometric | `hyper` | `m, n, k`  
log-normal   | `lnorm`   | `meanlog, sdlog`
logistic     | `logis`   | `location, scale`
negative binomial  | `nbinom` | `size, prob`
normal       | `norm`    | `mean, sd`
Poisson      | `pois`    | `lambda`
signed rank  | `signrank` | `n`
Student’s t  | `t`        | `df, ncp`
uniform      | `unif`     | `min, max`
Weibull      | `weibull`  | `shape, scale`
Wilcoxon     | `wilcox`   | `m, n`

## Exercise

### Normal

```{r norm6, fig.cap="Standard normal distribution", echo=FALSE}

set.seed(123)
data <- data.frame(x= c(rnorm(1000000,2,1)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    xlim(-2, 6)+
  ylab("f(x)")+
  ylim(0, 0.43)+
  geom_vline(xintercept = 2, colour ="red") +
  ggtitle("N(1,2)")

print(p)
```


```{r, eval = FALSE}
# ?dnorm - Help page

dnorm(2, mean = 2, sd = 1, log = FALSE)
pnorm(2, mean = 2, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(0.5, mean = 2, sd = 1, lower.tail = TRUE, log.p = FALSE)
rnorm(n=10, mean = 2, sd = 1)
```


```{r}
# ?dnorm - Help page

dnorm(2, mean = 2, sd = 1, log = FALSE)
pnorm(2, mean = 2, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(0.5, mean = 2, sd = 1, lower.tail = TRUE, log.p = FALSE)
rnorm(n=10, mean = 2, sd = 1)
```


### Gamma

```{r gamma, fig.cap="Standard normal distribution", echo=FALSE}

set.seed(123)
data <- data.frame(x= c(rgamma(1000000,shape = 2, scale =1/3)) )

dat <- with(density(data$x), data.frame(x, y))
p<- ggplot(data = dat, mapping = aes(x = x, y = y)) +
    geom_line()+
    xlim(0, 6)+
  ylab("f(x)")+
  ylim(0, 1.5)+
  ggtitle("gamma(2,3)")+geom_vline(xintercept = 1, color = "blue")

print(p)
```



```{r eval = FALSE}
dgamma(1, shape =2, scale = 1/3, log = FALSE)
pgamma(1, shape =2, scale = 1/3, lower.tail = TRUE,
       log.p = FALSE)
qgamma(0.8, shape = 2,  scale = 1/3, lower.tail = TRUE,
       log.p = FALSE)
rgamma(10, shape =2 , scale = 1/3)

```

 
```{r}
dgamma(1, shape =2, scale = 1/3, log = FALSE)
pgamma(1, shape =2, scale = 1/3, lower.tail = TRUE,
       log.p = FALSE)
qgamma(0.8, shape = 2,  scale = 1/3, lower.tail = TRUE,
       log.p = FALSE)
rgamma(10, shape =2 , scale = 1/3)

```

\newpage

## References

Discrete 
http://www.r-tutor.com/elementary-statistics/probability-distributions

https://rstudio-pubs-static.s3.amazonaws.com/100906_8e3a32dd11c14b839468db756cee7400.html